IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-03-02 paper/retrofit.tex
first subsection

We have typeset nonterminals of the target language using \textbf{{\color{red}
bold font}} while the nonterminals of the IFC language have been typeset
with \textit{{\color{blue} italic font}}.  Readers are encouraged to view
a color copy of this paper, where target language nonterminals are colored \textbf{{\color{red} red}}
while IFC language nonterminals are colored \textit{{\color{blue} blue}}.

@@

In Figure~\ref{fig:ml}, we give a simple, untyped lambda calculus with
mutable references and fixpoint, prepared for combination with an
information flow control language.  The presentation is standard, and utilizes Felleisen-Hieb reduction
semantics~\cite{Felleisen:1992:RRS:136293.136297} to define the
operational semantics of the system.  Our language defines an evaluation
context |tE|; however, the evaluation rules have been expressed in terms
of a different evaluation context |curlyRedE|,
following the approach of Matthews and
Findler~\cite{Matthews:2007:OSM:1190216.1190220} to combine operational
semantics of multiple languages.  To
derive the usual operational semantics for this language, the evaluation
context merely needs to be defined as |conf tS te @= tS, bracket tE te|.
However, when we combine this language with an IFC language, we must
reinterpret the meaning of this evaluation context.
In general, we require that a target language be expressed in terms
of some global machine state |tS|, some evaluation context |tE|,
some expressions |te|, some set of values |tv| and a \emph{deterministic}
reduction relation on full configurations $|tS| \times |tE| \times |te|$.

@@

Most modern, dynamic information flow control languages encode
policy by associating a \emph{label}
with data, which encodes whether or not data tagged with one
label |l1| can flow to another label |l2| (written as |canFlowTo2 l1 l2|).
Our embedding is agnostic to the choice of labeling scheme; we only
require the labels to form a lattice~\cite{Denning:1976:LMS:360051.360056}
with the partial order |canFlowTo|.
In this paper, we simply represent labels with the metavariable
$l$, but do not discuss them in more detail.

@@

\begin{itemize}
    \item First-class labels, with label values $l$ as well as operations for computing on
labels (|canFlowTo|, |lub| and |glb|),
    \item Operations for inspecting (\textbf{getLabel}) and modifying (\textbf{setLabel}) the current label of the task (a task can only increase its label, and not decrease it),
    \item Operations for non-blocking inter-task communication (\textbf{send} and \textbf{recv}),
        which interact with the global store of per-task message queues |iS|, and
    \item A sandboxing operation allowing to spawn new
    isolated tasks, which will be described in more detail in the
    next section.
\end{itemize}

While the first two bullet points are essentially universal to all
floating-label systems, our choice of communication primitives is motivated
by inter-worker communication in web browser APIs~\cite{webworkers}, but
other choices such as blocking communication or labeled channels are possible.
We describe some of the alternate possibilities in Section~\ref{sec:extensions}.
These operations are all defined with respect to an evaluation context |curlyBlueE| represents the context
of the current task which is running, which has three important pieces
of state: the global message queues |iS|,
the current label |il| (which protects the program counter of the task)
and the task ID |id|.

Our asynchronous communication primitives are worth some further
discussion.  When a task is sending a message using |send|, it also labels that
message with a label |il'| (which must be above the tasks current label |il|).
Messages can only be received by a task if its current label is
at least as high as the label of the message.
Specifically, receiving a message using
\[ |recvf ix1 ix2 ie1 ie2| \]
binds the message and the senders task identifier
to local variables |ix1| and |ix2|, respectively, and then executes |ie1|.
Otherwise the message gets removed from the queue by rule \textsc{I-noRecv},
and that tasks continues its execution with |ie2|.
We denote the filtering of the message queue by |filter il Q|,
which is defined as follows.
If |Q| is the empty list |nil|, the
function is simply the identity function, i.e.,
|filter il nil = nil|; otherwise:
\[
|filter il ((il', id, ie) , Q )| = \left\{
\begin{array}{l l}
|(il', id, ie) , filter il Q| & \quad \text{if |il' canFlowTo il|}\\
|filter il Q| & \quad \text{otherwise}
\end{array} \right.
\]
This ensures that tasks cannot receive messages that are more secret
than their current label would allow.

As a final remark, blocking receive can be (inefficiently) implemented using
our asynchronous primitives as follows:
\begin{code}
  brecvf ix1 ix2 ie @=
  tab tab tab (IT (fix (\ k . TI (recvf ix1 ix2 ie (IT k)) )))
\end{code}

@@

We note that it is possible to add rules for a natuaral embedding of
expressions that occur in both languages.  In our specific example
with |targetLangML|, we can choose to preserve booleans across
language boundaries as follows:
\begin{mathpar}
\inferrule[I-bool]
{ |b member {true, false}|  } {| iniE (IT b) -> iniE b |}
\and
\inferrule[T-bool]
{ |b member {true, false}| } {| intE (TI b) -> intE b |}
\end{mathpar}

>>>>

\paragraph{Notation}
We have typeset nonterminals of the target language using \textbf{{\color{red}
bold font}} while the nonterminals of the IFC language have been typeset
with \textit{{\color{blue} italic font}}.  Readers are encouraged to view
a color copy of this paper, where target language nonterminals are colored \textbf{{\color{red} red}}
while IFC language nonterminals are colored \textit{{\color{blue} blue}}.

@@

In Figure~\ref{fig:ml}, we give a simple, untyped lambda calculus with
mutable references and fixpoint, prepared for combination with an
information flow control language.  The presentation is standard, and utilizes Felleisen-Hieb reduction
semantics~\cite{Felleisen:1992:RRS:136293.136297} to define the
operational semantics of the system.  Our language defines an evaluation
context |tE|; however, the evaluation rules have been expressed in terms
of a different evaluation context |curlyRedE|,
following the approach of Matthews and
Findler~\cite{Matthews:2007:OSM:1190216.1190220} to simplify combining
semantics of multiple languages.  To
derive the usual operational semantics for this language, the evaluation
context merely needs to be defined as |conf tS te @= tS, bracket tE te|.
However, when we combine this language with an IFC language, we
reinterpret the meaning of this evaluation context.
In general, we require that a target language be expressed in terms
of some global machine state |tS|, some evaluation context |tE|,
some expressions |te|, some set of values |tv| and a \emph{deterministic}
reduction relation on full configurations $|tS| \times |tE| \times |te|$.

@@

Most modern, dynamic information flow control languages encode
policy by associating a \emph{label}
with data, which encodes whether or not data tagged with one
label |l1| can flow to another label |l2| (written as |canFlowTo2 l1 l2|).
Our embedding is agnostic to the choice of labeling scheme; we only
require the labels to form a lattice~\cite{Denning:1976:LMS:360051.360056}
with the partial order |canFlowTo|, join |lub|, and meet |glb|.
In this paper, we simply represent labels with the metavariable
$l$, but do not discuss them in more detail.

@@

\begin{itemize}
    \item First-class labels, with label values $l$ as well as operations for computing on
labels (|canFlowTo|, |lub| and |glb|),
    \item Operations for inspecting (\textbf{getLabel}) and modifying (\textbf{setLabel}) the current label of the task (a task can only increase its label, and not decrease it),
    \item Operations for non-blocking inter-task communication (\textbf{send} and \textbf{recv}),
        which interact with the global store of per-task message queues |iS|, and
    \item A sandboxing operation allowing to spawn new
    isolated tasks, described in more detail in the
    next section.
\end{itemize}

These operations are all defined with respect to an evaluation context
|curlyBlueE| represents the context of the current task which is running, which
has three important pieces of state: the global message queues |iS|, the
current label |il| and the task ID |id|.

At a high level, the semantics define a coarse-grained floating-label IFC
system~\cite{Zeldovich:2006, lio}, i.e., a system in which the current label of
the task can be raised to protect everything in scope.
%
We note that while first class labels and operations for inspecting the current
label are are essentially universal to all floating-label systems, our choice
of communication and isolation primitives is motivated by those present in
browsers, namely \texttt{postMessage} and Workers~\cite{webworkers,webmessaging}.
%
Of course, other choices such as blocking communication or labeled channels are
possible and further discussed in Section~\ref{sec:extensions}.
%
Though motivated by Web Workers, our notion of a task generalizes previous
primitives considered by floating-label dynamic IFC
systems~\cite{Zeldovich:2006, lio, stefan:addressing-covert,
Hritcu:2013:YIB:2497621.2498098}.

Our asynchronous communication primitives are worth some further
discussion.  When a task is sending a message using |send|, it also labels that
message with a label |il'| (which must be above the tasks current label |il|).
Messages can only be received by a task if its current label is
at least as high as the label of the message.
Specifically, receiving a message using
\[ |recvf ix1 ix2 ie1 ie2| \]
binds the message and the senders task identifier
to local variables |ix1| and |ix2|, respectively, and then executes |ie1|.
Otherwise the message gets removed from the queue by rule \textsc{I-noRecv},
and that tasks continues its execution with |ie2|.
We denote the filtering of the message queue by |filter il Q|,
which is defined as follows.
If |Q| is the empty list |nil|, the
function is simply the identity function, i.e.,
|filter il nil = nil|; otherwise:
\[
|filter il ((il', id, ie) , Q )| = \left\{
\begin{array}{l l}
|(il', id, ie) , filter il Q| & \quad \text{if |il' canFlowTo il|}\\
|filter il Q| & \quad \text{otherwise}
\end{array} \right.
\]
This ensures that tasks cannot receive messages that are more sensitive
than their current label would allow.

\Red{MOVE:}
As a final remark, blocking receive can be (inefficiently) implemented using
our asynchronous primitives as follows:
\begin{code}
  brecvf ix1 ix2 ie @=
  tab tab tab (IT (fix (\ k . TI (recvf ix1 ix2 ie (IT k)) )))
\end{code}

@@

We note that it is possible to add rules for a natural embedding of
expressions that occur in both languages.  In our specific example
with |targetLangML|, we can choose to preserve booleans across
language boundaries as follows:
\begin{mathpar}
\inferrule[I-bool]
{ |b member {true, false}|  } {| iniE (IT b) -> iniE b |}
\and
\inferrule[T-bool]
{ |b member {true, false}| } {| intE (TI b) -> intE b |}
\end{mathpar}
